---
title: "Type A Aortic Dissection"
subtitle: "NIS data analysis"
author: "Salil Deo "
date: "`r Sys.Date()`"
output: tint::tintHtml

link-citations: yes
---

```{r setup, include=FALSE}
library(tint)

```


this a document for analysis of only type a aortic dissection patients undergoing surgical repair. the dataframe is subsetted from the entire 2008 - 2015 data on aortic dissection created earlier. the original dataframe can be used to determine overall proportions if needed.

```{r load libraries}

library(haven);library(tidyverse);library(mosaic);library(survey)
library(skimr);library(srvyr);library(forcats)

options(survey.lonely.psu = "certainty")



```

load the dataframe

```{r load dataframe}

ta <- read_csv("D:/aortic_dissection/ta.csv")

skim(ta)


```

Survey methods for type a aortic dissection repair patients:

```{r}



# new design object

ta_s <- ta %>% srvyr::as_survey_design(weights = discwt.x, strata = nis_stratum.x, id = hosp_nis, nest = TRUE)


ta_s %>%
  summarise(age = survey_mean(age,NA.RM = TRUE, vartype = c("ci"), df = Inf))


ta_s %>%
  group_by(year.x) %>%
  summarise(age = survey_mean(age, na.rm = TRUE, vartype = c("ci"), df = Inf))


# gender overall

ta_s %>%
  summarise(female = survey_total(female, na.rm = TRUE, vartype = c("ci"), df = Inf))

# gender_year

ta_s %>%
  group_by(year.x) %>%
  summarise(female = survey_total(female, na.rm = TRUE, vartype = c("ci"), df = Inf))

# type of hospital

# overall for all type a aortic dissection repairs

survey::svytable(~hosp_type, design = ta_s)

# hospital type by year 

survey::svytable(~hosp_type + year.x, design = ta_s)

# race overall and by year

survey::svytable(~race_n, design = ta_s)

# insurance type overall and by year

survey::svytable(~insu,  design = ta_s)

survey::svytable(~year.x + insu,  design = ta_s)

# hospital size overall and by year

survey::svytable(~hosp_size, design = ta_s)

survey::svytable(~year.x + hosp_size, design = ta_s)


# renal failure patients overall and by year

ta_s %>%
  summarise(prop = survey_mean(cm_renlfail, na.rm = TRUE, vartype = c("ci"), deff = Inf))

# renal dysfunction by year

 ta_s %>%
  group_by(year.x) %>%
  summarise(prop = survey_mean(cm_renlfail, na.rm = TRUE, vartype = c("ci")))





# chronic lung disease/ COPD

ta_s %>%
  summarise(prop = survey_mean(cm_chrnlung, na.rm = TRUE, vartype = c("ci")))

 ta_s %>%
  group_by(year.x) %>%
  summarise(prop = survey_mean(cm_chrnlung, na.rm = TRUE, vartype = c("ci")))





# congestive heart failure 

ta_s %>%
  summarise(prop = survey_mean(cm_chf, na.rm = TRUE, vartype = c("ci")))

ta_s %>%
  group_by(year.x) %>%
  summarise(prop = survey_mean(cm_chf, na.rm = TRUE, vartype = c("ci")))



# specific etiology for aortic dissection
# marfan syndrome

ta_s %>%
  summarise(prop = survey_mean(marfan, na.rm = TRUE, vartype = c("ci")))

# ehler danlos syndrome

ta_s %>%
  summarise(prop = survey_mean(ehd, na.rm = TRUE, vartype = c("ci")))


# aortic ectasia

ta_s %>%
  summarise(prop = survey_mean(aae, na.rm = TRUE, vartype = c("ci")))


# uncomplicated and complicated hypertension

ta_s %>%
  summarise(prop = survey_mean(cm_htn_c, na.rm = TRUE, vartype = c("ci")))

ta_s %>%
  group_by(year.x) %>%
  summarise(prop = survey_mean(cm_htn_c, na.rm = TRUE, vartype = c("ci")))

# complicated dm patients

ta_s %>%
  summarise(prop = survey_mean(cm_dmcx, na.rm = TRUE, vartype = c("ci")))

ta_s %>%
  group_by(year.x) %>%
  summarise(prop = survey_mean(cm_dmcx, na.rm = TRUE, vartype = c("ci")))


# prior cardiac surgery

# icd9 code V45.89

prior_cs <- as.character(c("V4589"))

ta$prior_cs <- with(ta, ifelse((dx1 %in% prior_cs | dx2 %in% prior_cs | dx3 %in% prior_cs | dx4 %in% prior_cs | dx5 %in% prior_cs | dx7 %in% prior_cs | dx8 %in% prior_cs | dx9 %in% prior_cs | dx10 %in% prior_cs), 1, 0))

ta %>% count(prior_cs)


ta_s <- ta %>% srvyr::as_survey_design(weights = discwt.x, strata = nis_stratum.x, id = hosp_nis, nest = TRUE)

ta_s %>% 
  summarise(prop = survey_mean(prior_cs, na.rm = TRUE, vartype = c("ci")))

ta_s %>%
  group_by(year.x) %>%
  summarise(prop = survey_mean(prior_cs, na.rm = TRUE, vartype = c("ci")))


```

also create variables for postoperative events and outcome:


```{r}
# create variables for postoperative events:

#acute renal failure

aki <- as.character(c("5849"))

ta$aki <- with(ta, ifelse((dx1 %in% aki | dx2 %in% aki | dx3 %in% aki | dx4 %in% aki | dx5 %in% aki | dx7 %in% aki | dx8 %in% aki | dx9 %in% aki | dx10 %in% aki), 1, 0))

ta %>% count(aki)


# acute respiratory failure 
# icd9 code 51881

avi <- as.character(c("51881"))

ta$avi <- with(ta, ifelse((dx1 %in% avi | dx2 %in% avi | dx3 %in% avi | dx4 %in% avi | dx5 %in% avi | dx7 %in% avi | dx8 %in% avi | dx9 %in% avi | dx10 %in% avi), 1, 0))

ta %>% count(avi)

# acute postoperative stroke
# icd 9 code 43411

st <- as.character(c("43411"))

ta$stroke <- with(ta, ifelse((dx1 %in% st | dx2 %in% st | dx3 %in% st | dx4 %in% st | dx5 %in% st | dx7 %in% st | dx8 %in% st | dx9 %in% st | dx10 %in% st), 1, 0))

ta %>% count(stroke)

# acute abdomen / acute abdominal emergency
# icd9 code 789.00


acab <- as.character(c("78900"))

ta$acab <- with(ta, ifelse((dx1 %in% acab | dx2 %in% acab | dx3 %in% acab | dx4 %in% acab | dx5 %in% acab | dx7 %in% acab | dx8 %in% acab | dx9 %in% acab | dx10 %in% acab), 1, 0))

ta %>% count(acab)

# cardiac arrest
# icd 9 code 4275


c_arrest <- as.character(c("4275"))

ta$c_arrest <- with(ta, ifelse((dx1 %in% c_arrest | dx2 %in% c_arrest | dx3 %in% c_arrest | dx4 %in% c_arrest | dx5 %in% c_arrest | dx7 %in% c_arrest | dx8 %in% c_arrest | dx9 %in% c_arrest | dx10 %in% c_arrest), 1, 0))

ta %>% count(c_arrest)

# cardiogenic shock 
# icd 9 code 785.51


c_shock <- as.character(c("4275"))

ta$c_shock <- with(ta, ifelse((dx1 %in% c_shock | dx2 %in% c_shock | dx3 %in% c_shock | dx4 %in% c_shock | dx5 %in% c_shock | dx7 %in% c_shock | dx8 %in% c_shock | dx9 %in% c_shock | dx10 %in% c_shock), 1, 0))

ta %>% count(c_shock)



```

now plan for survey analysis...

```{r}

# redo survey object with the added variables

ta_s <- ta %>% srvyr::as_survey_design(weights = discwt.x, strata = nis_stratum.x, id = hosp_nis, nest = TRUE)

# mortality

ta_s %>%
  summarise(died = survey_mean(died, vartype = c("ci"), na.rm = TRUE))

# mortality according to year

ta_s %>%
  group_by(year.x) %>%
  summarise(died = survey_mean(died, vartype = c("ci"), na.rm = TRUE))

ta_s %>%
  group_by(year.x, hosp_type) %>%
  summarise(died = survey_mean(died, vartype = c("ci"), na.rm = TRUE))

# los total

ta_s %>%
  summarise(los = survey_mean(los, vartype = c("ci"), na.rm = TRUE))

# los according to year

ta_s %>%
  group_by(year.x) %>%
  summarise(los = survey_mean(los, vartype = c("ci"), na.rm = TRUE))

## clinical events during hospital admission

# acute kidney failure

ta_s %>%
  summarise(aki = survey_mean(aki, vartype = c("ci"), na.rm = TRUE))


# acute kidney failure per year

ta_s %>%
  group_by(year.x) %>%
  summarise(aki = survey_mean(aki, vartype = c("ci"), na.rm = TRUE))

# acute respiratory failure

ta_s %>%
  summarise(avi = survey_mean(avi, vartype = c("ci"), na.rm = T))

ta_s %>%
  group_by(year.x) %>%
  summarise(avi = survey_mean(avi, vartype = c("ci"), na.rm = T))

# cardiogenic shock/cardiac arrest overall and per year

ta_s %>%
  summarise(shock = survey_mean(c_shock, vartype = c("ci"), na.rm = T))
```

create table one according to first and last of study:

```{r table one}

tabone <- ta %>%
  filter(year.x == 2008 | year.x == 2015)


```

